Clinical prediction with localized modeling using similarity-based cohorts: A scoping review

### 2023-09-10 docs/Manuscript/Manuscript.Rmd
make final rmarkdown revisions

The `r nrow(tab_synthesis)` included studies were analyzed based on 20 characteristics (see Section \ref{sec:data}), and we report and describe some observations here (Table \ref{tab:synthesis}). The studies were published in a variety of journals, with some of greater frequency, though no single journal published more than 3. The journal _Evolving Systems_ published 2 of the included studies, _Artificial Intelligence in Medicine_ published 3, _Hindawi - Journal of Healthcare Engineering_ published 3, and _PLOS One_ published 2.

@@

Less common were calls to strengthen the connection between the methods and the users.
Some authors urged the incorporation of (human-derived) domain knowledge into the data or models [@Yearwood1997a; @Wang2019]. Others suggested new uses of their modeling approaches: to measure feature importance [@Wyns2004a], to make models more expressive [@Lee2015], and to combine information obtained both from global and from localized models [@Ng2015].
Most of this incremental work was indeed carried out in later investigations in our sample.

## Reproducibility
\label{sec:reproducibility}

**Discuss public availability of code or implementations.**
**Methods, results, and inferential reproducibility [@Matarese2022].**

@@

Because we excluded studies that used conventional NN prediction, many included studies reported new approaches to adaptation subsequent to retrieval. Very few proposed novel similarity measures, possibly because larger studies tended to be reported in separate articles detailing experiments with specific components of the process. However, this also suggests that few experimental studies have focused on the unified development of new retrieval and adaptation strategies.
We also note that the majority of studies evaluated and compared methods using leave-one-out cross-validation (LOOCV). This is an especially appropriate technique when available data are scarce, and indeed most data sets used by included studies numbered in the hundreds of cases or fewer. This likely follows from the older age of many included studies and from their consistent primary focus on clinical data, which is more costly to collect and comes with more restrictions on its use.

@@

As discussed in Section \ref{sec:rationales}, most studies were premised on the potential predictive value of localized models.
As cautioned in Section \ref{sec:performance}, not all experiments affirmed this premise. Most of those that did involved using metric learning to improve retrieval and sometimes iterative optimization of the metric and of the localized model.
Some studies showcased results using a variety of such specializations [@CampilloGimenez2012a; @Ng2015; @Zhang2018a]. However, because these advanced implementations were mostly compared against their precursors or commonplace alternatives, we cannot speak to their performance or any trade-offs with respect to each other.
This would be a worthy goal of future work, though it might be limited by the lack of open-source code or public implementations (Section \ref{sec:reproducibility}).

While most studies identified outstanding technical needs (Section \ref{sec:needs}), few emphasized the need to assess human-focused qualities like user interface and user experience, interpretability, meaningfulness, or trust.
Most studies also identified interpretability as an advantage of localized models, though none evaluated model interpretability and few proposed new interpretative uses.
Lack of trust in ML tools is a long-recognized problem that direct intepretability of model components could help alleviate, but this is more often assumed than demonstrated.
One valuable direction for future work would be to compare the ability of non-technical investigators or users to draw correct inferences from these and other interpretable methods and their levels of confidence in those inferences.

We posit that an essential component of such "soft" performance goals is the ability of the user community to exert some control over the models.
Given the impact on performance of the choice or optimization of the similarity measure, an important target for user input would be the importance of certain variables in the calculation of similarity, with an understanding of how it can impact not only the performance of the predictions but also the cohort retrieved for the model. For one example, a user may want to minimize the weight of rare diseases in medical history in order to retrieve a population with more such cases in order to better measure their associated risk to the outcome. In contrast, they may want to increase the weight of the indicating diagnosis in order to allow fewer patients from similar but distinct populations to influence the model. For another example, a user may want to down-weight socioeconomic variables like race--ethnicity in order to ensure a more diverse modeling cohort.
Future purely quantitative work could assess whether similarity-tuning can achieve these ends more flexibly than strict inclusion/exclusion criteria.

>>>>

The `r nrow(tab_synthesis)` included studies were analyzed based on several characteristics (see Section \ref{sec:data}), and we report and describe some observations here (Table \ref{tab:synthesis}). The studies were published in a variety of journals, with some of greater frequency, though no single journal published more than 3. The journal _Evolving Systems_ published 2 of the included studies, _Artificial Intelligence in Medicine_ published 3, _Hindawi - Journal of Healthcare Engineering_ published 3, and _PLOS One_ published 2.

@@

Less common were calls to strengthen the connection between the methods and the users.
Some authors urged the incorporation of (human-derived) domain knowledge into the data or models [@Yearwood1997a; @Wang2019]. Others suggested new uses of their modeling approaches: to measure feature importance [@Wyns2004a], to make models more expressive [@Lee2015], and to combine information obtained both from global and from localized models [@Ng2015].
Most of this incremental work was indeed carried out in later investigations in our sample.
That said, we did not find reports of successful implementations of these tools in clinical practice. (Several excluded studies did showcase uses of CBR in practice.)

@@

Because we excluded studies that used conventional NN prediction, many included studies reported new approaches to adaptation subsequent to retrieval. Very few proposed novel similarity measures, possibly because larger studies tended to be reported in separate articles detailing experiments with specific components of the process. However, this also suggests that few experimental studies have focused on the unified development of new retrieval and adaptation strategies.
We also note that the majority of studies evaluated and compared methods using leave-one-out cross-validation (LOOCV). This is an appropriate technique when available data are scarce, and indeed most data sets used by included studies numbered in the hundreds of cases or fewer. This likely follows from the older age of many included studies and from their consistent primary focus on clinical data, which is more costly to collect and comes with more restrictions on its use.

@@

As discussed in Section \ref{sec:rationales}, most studies were premised on the potential predictive value of localized models.
As cautioned in Section \ref{sec:performance}, not all experiments affirmed this premise. Most of those that did involved using metric learning to improve retrieval and sometimes iterative optimization of the metric and of the localized model.
Some studies showcased results using a variety of such specializations [@CampilloGimenez2012a; @Ng2015; @Zhang2018a]. However, because these advanced implementations were mostly compared against their precursors or commonplace alternatives, we cannot speak to their performance or any trade-offs with respect to each other.
This would be a worthy goal of future work.

While most studies identified outstanding technical needs (Section \ref{sec:needs}), few emphasized the need to assess human-focused qualities like user interface and user experience, interpretability, meaningfulness, or trust.
Most studies identified interpretability as an advantage of localized models, though none evaluated model interpretability and few proposed new interpretative uses.
Lack of trust in ML tools is a long-recognized problem that direct intepretability of model components could help alleviate, but this is more often assumed than demonstrated.
One valuable direction for future work would be to measure the utility of interpretable components in research and in clinical practice and the correctness and confidence of users in their interpretations.

The primary goal of these studies was to establish that localized models, and certain strategies within this paradigm, perform at or above the level of other predictive models or strategies.
From a functionalist approach to reproducibility [@Matarese2022], then, later studies were broadly successful at reproducing earlier studies, and later studies provide sufficient detail to support ongoing reproduction efforts---though these might be limited by the lack of open-source code or public implementations.
However, the immediate goal of most studes was to improve care, and an implied need, often enjoined but neither obtained nor reproduced, was a demonstration of practical use.

We posit that an essential component of practical usefulness is the ability of the user community to exert some control over the models.
Given the impact on performance of the choice or optimization of the similarity measure, an important target for user input would be the importance of certain variables in the calculation of similarity, with an understanding of how it can impact not only the performance of the predictions but also the cohort retrieved for the model. For one example, a user may want to minimize the weight of rare diseases in medical history in order to retrieve a population with more such cases in order to better measure their associated risk to the outcome. In contrast, they may want to increase the weight of the indicating diagnosis in order to allow fewer patients from similar but distinct populations to influence the model. For another example, a user may want to down-weight socioeconomic variables like race--ethnicity in order to ensure a more diverse modeling cohort.
Future purely quantitative work could assess whether similarity-tuning can achieve these ends more flexibly than strict inclusion/exclusion criteria.

### 2023-09-10 docs/Manuscript/Manuscript.tex
make final rmarkdown revisions

The 25 included studies were analyzed based on 20 characteristics (see
Section \ref{sec:data}), and we report and describe some observations
here (Table \ref{tab:synthesis}). The studies were published in a
variety of journals, with some of greater frequency, though no single
journal published more than 3. The journal \emph{Evolving Systems}
published 2 of the included studies, \emph{Artificial Intelligence in
Medicine} published 3, \emph{Hindawi - Journal of Healthcare
Engineering} published 3, and \emph{PLOS One} published 2.

@@

Less common were calls to strengthen the connection between the methods
and the users. Some authors urged the incorporation of (human-derived)
domain knowledge into the data or models {[}17,19{]}. Others suggested
new uses of their modeling approaches: to measure feature importance
{[}20{]}, to make models more expressive {[}14{]}, and to combine
information obtained both from global and from localized models
{[}15{]}. Most of this incremental work was indeed carried out in later
investigations in our sample.

\hypertarget{reproducibility}{%
\subsection{Reproducibility}\label{reproducibility}}

\label{sec:reproducibility}

\textbf{Discuss public availability of code or implementations.}
\textbf{Methods, results, and inferential reproducibility {[}39{]}.}

@@

These programs used varying terminology for common concepts, and no
common term is in use for what we here refer to as localized modeling;
authors described their approaches as ``targeted prognosis'' {[}11{]},
``transductive inference'' {[}21{]}, ``personalized decision support''
{[}14{]}, ``personalized (predictive) models'' {[}26, in contrast to
``local models,''15,17,32{]}, and ``precision cohort'' analysis
{[}17,34,35{]}. We find uses of ``targeted'', ``personalized'', and
``precision'' generic and imprecise to this approach, while transductive
inference is an established term for a broader set of methods in ML.
Because \emph{local} naturally contrasts with \emph{global}, we propose
``localized models'' as a suitable term for this counterpart to
globally-fitted models.\footnote{We note, in response to one reviewer's
  comment, that for many if not most models used in ML, excepting GLMs,
  a predictor may play a different and more or less important role for
  some cases than others, as quantified by importance measures or
  model-agnostic explanations {[}40,41{]}. In this way, such models can
  also be viewed as ``localized''. We recognize that the use of global
  and local explanations is standard. As another reviewer pointed out,
  the terminology of local and global is also used in the unrelated
  context of federated ML to distinguish models trained using data
  stored in a single storage device (local) from their aggregations
  (global) {[}42,43{]}. The prospects for federated architecture to more
  effectively implement similarity cohort--based models are intriguing,
  and would require some reconciliation of terms. At present, we feel
  that these methods are sufficiently disjoint to avoid confusion in
  practice, but we suggest the more specific term ``similarity
  cohort--based models'' when the term ``local'' is overloaded.}

@@

The 25 studies included in our synthesis used composite techniques that
we organized into two major types. One type, \emph{similarity learning},
was used in 8 studies that tie in to the much larger literature on DML
{[}44{]}. The other, \emph{cohort thresholding}, focused on choosing or
optimizing the manner in which cohorts were demarcated using the
similarity measure.

@@

The similarity learning approaches took a variety of forms. These
techniques are designed to detect structure in data, especially
associations between predictors and responses, and to use this structure
to inform the definition of a patient similarity measure. Seven studies
used response values in training data sets to supervise similarity
learning while two used unsupervised learning (Table
\ref{tab:composite}). Based on the key properties of DML algorithms
identified by {[}45{]}, most learned measures were non-linear while some
were locally learned, and some were optimized globally while others
locally. Not all qualified as DML, since the resulting measures would
not necessarily satisfy the triangle inequality. Each used its
similarity measure to retrieve training cases relevant to each testing
case. Once retrieved, most took a standard NN approach to generating
predictions; two exceptions {[}29,34{]} fit predictive models to the
retrieved cohorts.

Several studies used unsupervised similarity learning, for example the
Mahalanobis distance {[}12{]} and the Kolmogorov entropy-based distance
{[}22{]}. In particular, {[}19{]} defined similarity as a weighted sum
of differences in predictor values and used linear regression to
optimize the weights for the predictive accuracy of the cohort they
retrieve. Others used supervised learning: {[}21{]} proposed an
iterative algorithm to optimize a set of fuzzy inference rules used to
retrieve relevant cases and generate a prediction, using
back-propagation on the rules' parameters. Their method was used in
several later studies returned by our search, which are not discussed
here because they did not originate the technique. {[}28{]} explicitly
appeal to a supervised distance metric learning (DML) technique {[}46{]}
to obtain a similarity measure on their set of dermoscopy and confocal
images that most effectively separates malignant from benign melanoma
tumors. We will discuss the remaining uses of supervised similarity
learning shortly.

Distinct from but related to similarity learning was supervised cohort
construction. This was an adaptive step taken by {[}13{]} to optimize
the predictive performance of models fitted to cohorts retrieved using a
fixed similarity rather than count threshold, a counterpart to
conventional CBR they called ``statistical CBR''.\footnote{Drawing from
  {[}47{]}, we suggest the term ``combinatorial CBR'' for the
  then-conventional approach using similarity cohorts of fixed
  cardinality.} The step used a heuristic procedure to locate a
similarity threshold that (locally) maximizes predictive accuracy on the
training set. This is analogous to optimizing the neighborhood size
parameter in NN predictive modeling, so we will not discuss it further
except to mention that the authors found statistical CBR to outperform
conventional CBR on several data sets. {[}27{]} combined these learning
techniques: They used logistic regression on the training set to obtain
weights for the predictors (by predicting the binary outcome of kidney
transplant waitlist registration) and for the cases (by predicting
agreement of outcome between an index case and other training cases),
and they used exhaustion to optimize the size of the retrieved cohort
for the accuracy of the NN model using the previously optimized weights.

@@

As discussed in Section \ref{sec:rationales}, most studies were premised
on the potential predictive value of localized models. As cautioned in
Section \ref{sec:performance}, not all experiments affirmed this
premise. Most of those that did involved using metric learning to
improve retrieval and sometimes iterative optimization of the metric and
of the localized model. Some studies showcased results using a variety
of such specializations {[}15,27,30{]}. However, because these advanced
implementations were mostly compared against their precursors or
commonplace alternatives, we cannot speak to their performance or any
trade-offs with respect to each other. This would be a worthy goal of
future work, though it might be limited by the lack of open-source code
or public implementations (Section \ref{sec:reproducibility}).

While most studies identified outstanding technical needs (Section
\ref{sec:needs}), few emphasized the need to assess human-focused
qualities like user interface and user experience, interpretability,
meaningfulness, or trust. Most studies also identified interpretability
as an advantage of localized models, though none evaluated model
interpretability and few proposed new interpretative uses. Lack of trust
in ML tools is a long-recognized problem that direct intepretability of
model components could help alleviate, but this is more often assumed
than demonstrated. One valuable direction for future work would be to
compare the ability of non-technical investigators or users to draw
correct inferences from these and other interpretable methods and their
levels of confidence in those inferences.

We posit that an essential component of such ``soft'' performance goals
is the ability of the user community to exert some control over the
models. Given the impact on performance of the choice or optimization of
the similarity measure, an important target for user input would be the
importance of certain variables in the calculation of similarity, with
an understanding of how it can impact not only the performance of the
predictions but also the cohort retrieved for the model. For one
example, a user may want to minimize the weight of rare diseases in
medical history in order to retrieve a population with more such cases
in order to better measure their associated risk to the outcome. In
contrast, they may want to increase the weight of the indicating
diagnosis in order to allow fewer patients from similar but distinct
populations to influence the model. For another example, a user may want
to down-weight socioeconomic variables like race--ethnicity in order to
ensure a more diverse modeling cohort. Future purely quantitative work
could assess whether similarity-tuning can achieve these ends more
flexibly than strict inclusion/exclusion criteria.

@@

\leavevmode\vadjust pre{\hypertarget{ref-Matarese2022}{}}%
\CSLLeftMargin{{[}39{]} }%
\CSLRightInline{Matarese V. Kinds of {Replicability}: {Different Terms}
and {Different Functions}. Axiomathes 2022;32:647--70.
\url{https://doi.org/10.1007/s10516-021-09610-2}.}

\leavevmode\vadjust pre{\hypertarget{ref-Biecek2021}{}}%
\CSLLeftMargin{{[}40{]} }%
\CSLRightInline{Biecek P, Burzykowski T. Explanatory {Model Analysis}:
{Explore}, {Explain}, and {Examine Predictive Models}. {New York}:
{Chapman and Hall/CRC}; 2021.}

\leavevmode\vadjust pre{\hypertarget{ref-Molnar2023}{}}%
\CSLLeftMargin{{[}41{]} }%
\CSLRightInline{Molnar C. Interpretable {Machine Learning}: {A Guide}
for {Making Black Box Models Explainable}. 2023.}

\leavevmode\vadjust pre{\hypertarget{ref-Moshawrab2023}{}}%
\CSLLeftMargin{{[}42{]} }%
\CSLRightInline{Moshawrab M, Adda M, Bouzouane A, Ibrahim H, Raad A.
Reviewing {Federated Machine Learning} and {Its Use} in {Diseases
Prediction}. Sensors 2023;23:2112.
\url{https://doi.org/10.3390/s23042112}.}

\leavevmode\vadjust pre{\hypertarget{ref-Brauneck2023}{}}%
\CSLLeftMargin{{[}43{]} }%
\CSLRightInline{Brauneck A, Schmalhorst L, Majdabadi MMK, Bakhtiari M,
Völker U, Baumbach J, et al. Federated {Machine Learning},
{Privacy-Enhancing Technologies}, and {Data Protection Laws} in {Medical
Research}: {Scoping Review}. Journal of Medical Internet Research
2023;25:e41588. \url{https://doi.org/10.2196/41588}.}

\leavevmode\vadjust pre{\hypertarget{ref-Yang2006}{}}%
\CSLLeftMargin{{[}44{]} }%
\CSLRightInline{Yang L. Distance {Metric Learning}: {A Comprehensive
Survey}. PhD thesis. Michigan State University, 2006.}

\leavevmode\vadjust pre{\hypertarget{ref-Bellet2014}{}}%
\CSLLeftMargin{{[}45{]} }%
\CSLRightInline{Bellet A, Habrard A, Sebban M. A {Survey} on {Metric
Learning} for {Feature Vectors} and {Structured Data} 2014.
\url{https://doi.org/10.48550/arXiv.1306.6709}.}

\leavevmode\vadjust pre{\hypertarget{ref-Xing2002}{}}%
\CSLLeftMargin{{[}46{]} }%
\CSLRightInline{Xing E, Jordan M, Russell SJ, Ng A. Distance {Metric
Learning} with {Application} to {Clustering} with {Side-Information}.
Advances in {Neural Information Processing Systems}, vol. 15, {MIT
Press}; 2002.}

\leavevmode\vadjust pre{\hypertarget{ref-Goyal2008}{}}%
\CSLLeftMargin{{[}47{]} }%
\CSLRightInline{Goyal N, Lifshits Y, Schütze H. Disorder inequality: A
combinatorial approach to nearest neighbor search. Proceedings of the
2008 {International Conference} on {Web Search} and {Data Mining}, {New
York, NY, USA}: {Association for Computing Machinery}; 2008, p. 25--32.
\url{https://doi.org/10.1145/1341531.1341538}.}

>>>>

The 25 included studies were analyzed based on several characteristics
(see Section \ref{sec:data}), and we report and describe some
observations here (Table \ref{tab:synthesis}). The studies were
published in a variety of journals, with some of greater frequency,
though no single journal published more than 3. The journal
\emph{Evolving Systems} published 2 of the included studies,
\emph{Artificial Intelligence in Medicine} published 3, \emph{Hindawi -
Journal of Healthcare Engineering} published 3, and \emph{PLOS One}
published 2.

@@

Less common were calls to strengthen the connection between the methods
and the users. Some authors urged the incorporation of (human-derived)
domain knowledge into the data or models {[}17,19{]}. Others suggested
new uses of their modeling approaches: to measure feature importance
{[}20{]}, to make models more expressive {[}14{]}, and to combine
information obtained both from global and from localized models
{[}15{]}. Most of this incremental work was indeed carried out in later
investigations in our sample. That said, we did not find reports of
successful implementations of these tools in clinical practice. (Several
excluded studies did showcase uses of CBR in practice.)

@@

These programs used varying terminology for common concepts, and no
common term is in use for what we here refer to as localized modeling;
authors described their approaches as ``targeted prognosis'' {[}11{]},
``transductive inference'' {[}21{]}, ``personalized decision support''
{[}14{]}, ``personalized (predictive) models'' {[}26, in contrast to
``local models,''15,17,32{]}, and ``precision cohort'' analysis
{[}17,34,35{]}. We find uses of ``targeted'', ``personalized'', and
``precision'' generic and imprecise to this approach, while transductive
inference is an established term for a broader set of methods in ML.
Because \emph{local} naturally contrasts with \emph{global}, we propose
``localized models'' as a suitable term for this counterpart to
globally-fitted models.\footnote{We note, in response to one reviewer's
  comment, that for many if not most models used in ML, excepting GLMs,
  a predictor may play a different and more or less important role for
  some cases than others, as quantified by importance measures or
  model-agnostic explanations {[}39,40{]}. In this way, such models can
  also be viewed as ``localized''. We recognize that the use of global
  and local explanations is standard. As another reviewer pointed out,
  the terminology of local and global is also used in the unrelated
  context of federated ML to distinguish models trained using data
  stored in a single storage device (local) from their aggregations
  (global) {[}41,42{]}. The prospects for federated architecture to more
  effectively implement similarity cohort--based models are intriguing,
  and would require some reconciliation of terms. At present, we feel
  that these methods are sufficiently disjoint to avoid confusion in
  practice, but we suggest the more specific term ``similarity
  cohort--based models'' when the term ``local'' is overloaded.}

@@

The 25 studies included in our synthesis used composite techniques that
we organized into two major types. One type, \emph{similarity learning},
was used in 8 studies that tie in to the much larger literature on DML
{[}43{]}. The other, \emph{cohort thresholding}, focused on choosing or
optimizing the manner in which cohorts were demarcated using the
similarity measure.

@@

The similarity learning approaches took a variety of forms. These
techniques are designed to detect structure in data, especially
associations between predictors and responses, and to use this structure
to inform the definition of a patient similarity measure. Seven studies
used response values in training data sets to supervise similarity
learning while two used unsupervised learning (Table
\ref{tab:composite}). Based on the key properties of DML algorithms
identified by {[}44{]}, most learned measures were non-linear while some
were locally learned, and some were optimized globally while others
locally. Not all qualified as DML, since the resulting measures would
not necessarily satisfy the triangle inequality. Each used its
similarity measure to retrieve training cases relevant to each testing
case. Once retrieved, most took a standard NN approach to generating
predictions; two exceptions {[}29,34{]} fit predictive models to the
retrieved cohorts.

Several studies used unsupervised similarity learning, for example the
Mahalanobis distance {[}12{]} and the Kolmogorov entropy-based distance
{[}22{]}. In particular, {[}19{]} defined similarity as a weighted sum
of differences in predictor values and used linear regression to
optimize the weights for the predictive accuracy of the cohort they
retrieve. Others used supervised learning: {[}21{]} proposed an
iterative algorithm to optimize a set of fuzzy inference rules used to
retrieve relevant cases and generate a prediction, using
back-propagation on the rules' parameters. Their method was used in
several later studies returned by our search, which are not discussed
here because they did not originate the technique. {[}28{]} explicitly
appeal to a supervised distance metric learning (DML) technique {[}45{]}
to obtain a similarity measure on their set of dermoscopy and confocal
images that most effectively separates malignant from benign melanoma
tumors. We will discuss the remaining uses of supervised similarity
learning shortly.

Distinct from but related to similarity learning was supervised cohort
construction. This was an adaptive step taken by {[}13{]} to optimize
the predictive performance of models fitted to cohorts retrieved using a
fixed similarity rather than count threshold, a counterpart to
conventional CBR they called ``statistical CBR''.\footnote{Drawing from
  {[}46{]}, we suggest the term ``combinatorial CBR'' for the
  then-conventional approach using similarity cohorts of fixed
  cardinality.} The step used a heuristic procedure to locate a
similarity threshold that (locally) maximizes predictive accuracy on the
training set. This is analogous to optimizing the neighborhood size
parameter in NN predictive modeling, so we will not discuss it further
except to mention that the authors found statistical CBR to outperform
conventional CBR on several data sets. {[}27{]} combined these learning
techniques: They used logistic regression on the training set to obtain
weights for the predictors (by predicting the binary outcome of kidney
transplant waitlist registration) and for the cases (by predicting
agreement of outcome between an index case and other training cases),
and they used exhaustion to optimize the size of the retrieved cohort
for the accuracy of the NN model using the previously optimized weights.

@@

As discussed in Section \ref{sec:rationales}, most studies were premised
on the potential predictive value of localized models. As cautioned in
Section \ref{sec:performance}, not all experiments affirmed this
premise. Most of those that did involved using metric learning to
improve retrieval and sometimes iterative optimization of the metric and
of the localized model. Some studies showcased results using a variety
of such specializations {[}15,27,30{]}. However, because these advanced
implementations were mostly compared against their precursors or
commonplace alternatives, we cannot speak to their performance or any
trade-offs with respect to each other. This would be a worthy goal of
future work.

While most studies identified outstanding technical needs (Section
\ref{sec:needs}), few emphasized the need to assess human-focused
qualities like user interface and user experience, interpretability,
meaningfulness, or trust. Most studies identified interpretability as an
advantage of localized models, though none evaluated model
interpretability and few proposed new interpretative uses. Lack of trust
in ML tools is a long-recognized problem that direct intepretability of
model components could help alleviate, but this is more often assumed
than demonstrated. One valuable direction for future work would be to
measure the utility of interpretable components in research and in
clinical practice and the correctness and confidence of users in their
interpretations.

The primary goal of these studies was to establish that localized
models, and certain strategies within this paradigm, perform at or above
the level of other predictive models or strategies. From a functionalist
approach to reproducibility {[}47{]}, then, later studies were broadly
successful at reproducing earlier studies, and later studies provide
sufficient detail to support ongoing reproduction efforts---though these
might be limited by the lack of open-source code or public
implementations. However, the immediate goal of most studes was to
improve care, and an implied need, often enjoined but neither obtained
nor reproduced, was a demonstration of practical use.

We posit that an essential component of practical usefulness is the
ability of the user community to exert some control over the models.
Given the impact on performance of the choice or optimization of the
similarity measure, an important target for user input would be the
importance of certain variables in the calculation of similarity, with
an understanding of how it can impact not only the performance of the
predictions but also the cohort retrieved for the model. For one
example, a user may want to minimize the weight of rare diseases in
medical history in order to retrieve a population with more such cases
in order to better measure their associated risk to the outcome. In
contrast, they may want to increase the weight of the indicating
diagnosis in order to allow fewer patients from similar but distinct
populations to influence the model. For another example, a user may want
to down-weight socioeconomic variables like race--ethnicity in order to
ensure a more diverse modeling cohort. Future purely quantitative work
could assess whether similarity-tuning can achieve these ends more
flexibly than strict inclusion/exclusion criteria.

@@

\leavevmode\vadjust pre{\hypertarget{ref-Biecek2021}{}}%
\CSLLeftMargin{{[}39{]} }%
\CSLRightInline{Biecek P, Burzykowski T. Explanatory {Model Analysis}:
{Explore}, {Explain}, and {Examine Predictive Models}. {New York}:
{Chapman and Hall/CRC}; 2021.}

\leavevmode\vadjust pre{\hypertarget{ref-Molnar2023}{}}%
\CSLLeftMargin{{[}40{]} }%
\CSLRightInline{Molnar C. Interpretable {Machine Learning}: {A Guide}
for {Making Black Box Models Explainable}. 2023.}

\leavevmode\vadjust pre{\hypertarget{ref-Moshawrab2023}{}}%
\CSLLeftMargin{{[}41{]} }%
\CSLRightInline{Moshawrab M, Adda M, Bouzouane A, Ibrahim H, Raad A.
Reviewing {Federated Machine Learning} and {Its Use} in {Diseases
Prediction}. Sensors 2023;23:2112.
\url{https://doi.org/10.3390/s23042112}.}

\leavevmode\vadjust pre{\hypertarget{ref-Brauneck2023}{}}%
\CSLLeftMargin{{[}42{]} }%
\CSLRightInline{Brauneck A, Schmalhorst L, Majdabadi MMK, Bakhtiari M,
Völker U, Baumbach J, et al. Federated {Machine Learning},
{Privacy-Enhancing Technologies}, and {Data Protection Laws} in {Medical
Research}: {Scoping Review}. Journal of Medical Internet Research
2023;25:e41588. \url{https://doi.org/10.2196/41588}.}

\leavevmode\vadjust pre{\hypertarget{ref-Yang2006}{}}%
\CSLLeftMargin{{[}43{]} }%
\CSLRightInline{Yang L. Distance {Metric Learning}: {A Comprehensive
Survey}. PhD thesis. Michigan State University, 2006.}

\leavevmode\vadjust pre{\hypertarget{ref-Bellet2014}{}}%
\CSLLeftMargin{{[}44{]} }%
\CSLRightInline{Bellet A, Habrard A, Sebban M. A {Survey} on {Metric
Learning} for {Feature Vectors} and {Structured Data} 2014.
\url{https://doi.org/10.48550/arXiv.1306.6709}.}

\leavevmode\vadjust pre{\hypertarget{ref-Xing2002}{}}%
\CSLLeftMargin{{[}45{]} }%
\CSLRightInline{Xing E, Jordan M, Russell SJ, Ng A. Distance {Metric
Learning} with {Application} to {Clustering} with {Side-Information}.
Advances in {Neural Information Processing Systems}, vol. 15, {MIT
Press}; 2002.}

\leavevmode\vadjust pre{\hypertarget{ref-Goyal2008}{}}%
\CSLLeftMargin{{[}46{]} }%
\CSLRightInline{Goyal N, Lifshits Y, Schütze H. Disorder inequality: A
combinatorial approach to nearest neighbor search. Proceedings of the
2008 {International Conference} on {Web Search} and {Data Mining}, {New
York, NY, USA}: {Association for Computing Machinery}; 2008, p. 25--32.
\url{https://doi.org/10.1145/1341531.1341538}.}

\leavevmode\vadjust pre{\hypertarget{ref-Matarese2022}{}}%
\CSLLeftMargin{{[}47{]} }%
\CSLRightInline{Matarese V. Kinds of {Replicability}: {Different Terms}
and {Different Functions}. Axiomathes 2022;32:647--70.
\url{https://doi.org/10.1007/s10516-021-09610-2}.}
